%% IntSeqBERT — CICM 2026
%% Main LaTeX file (LNCS format)
%% Compile: pdflatex main && bibtex main && pdflatex main && pdflatex main
%% -----------------------------------------------------------------------

\documentclass[runningheads]{llncs}

\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}  % for \toprule, \midrule, \bottomrule (optional)
\usepackage{url}
\usepackage{hyperref}
%\usepackage{microtype} % improve text justification (disabled: font expansion issue)

% If you use the hyperref package, uncomment below for Springer eBook style URLs:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}

%% -----------------------------------------------------------------------
%% Title and authors
%% -----------------------------------------------------------------------

\title{IntSeqBERT: Learning Arithmetic Structure in OEIS
       via Modulo-Spectrum Embeddings}

\titlerunning{IntSeqBERT: Learning Arithmetic Structure in OEIS}

\author{Kazuhisa Nakasho\inst{1}}

\authorrunning{K. Nakasho}

\institute{Iwate Prefectural University, Takizawa, Iwate, Japan\\
\email{nakasho\_k@iwate-pu.ac.jp}}

\maketitle

%% -----------------------------------------------------------------------
%% Abstract
%% -----------------------------------------------------------------------

\begin{abstract}
Integer sequences in the OEIS span values from single-digit constants to astronomical
factorials and exponentials, making prediction challenging for standard tokenised models that
cannot handle out-of-vocabulary values or exploit periodic arithmetic structure.
We present \textbf{IntSeqBERT}, a dual-stream Transformer encoder for masked integer-sequence
modelling on OEIS. Each sequence element is encoded along two complementary axes: a continuous
log-scale magnitude embedding and sin/cos modulo embeddings for 100 residues (moduli
$2$--$101$), fused via FiLM. Three prediction heads (magnitude regression,
sign classification, and modulo prediction for 100 moduli) are trained jointly
on 274,705 OEIS sequences.
At the Large scale (91.5M parameters), IntSeqBERT achieves 95.85\% magnitude accuracy and
50.38\% Mean Modulo Accuracy (MMA) on the test set, outperforming a standard tokenised
Transformer baseline by $+8.9$\,pt and $+4.5$\,pt, respectively. An ablation removing the
modulo stream confirms it accounts for $+15.2$\,pt of the MMA gain and contributes an
additional $+6.2$\,pt to magnitude accuracy. A probabilistic Chinese Remainder Theorem (CRT)-based Solver converts the model's
predictions into concrete integers, yielding a 7.4-fold improvement in next-term prediction
over the tokenised-Transformer baseline (Top-1: 19.09\% vs.\ 2.59\%). Modulo spectrum analysis reveals a strong negative correlation
between Normalised Information Gain (NIG) and Euler's totient ratio $\varphi(m)/m$
($r = -0.851$, $p < 10^{-28}$), providing empirical evidence that composite moduli capture
OEIS arithmetic structure more efficiently via CRT aggregation.

\keywords{Integer sequences \and OEIS \and Masked sequence modelling \and
          Modular arithmetic \and Transformer \and FiLM}
\end{abstract}

%% -----------------------------------------------------------------------
%% Sections
%% -----------------------------------------------------------------------

\input{sections/01_introduction}
\input{sections/02_background}
\input{sections/03_model}
\input{sections/04_setup}
\input{sections/05_experiments}
\input{sections/06_analysis}
\input{sections/07_conclusion}

%% -----------------------------------------------------------------------
%% Acknowledgements and Competing Interests
%% -----------------------------------------------------------------------

\begin{credits}
\subsubsection{\ackname}
This work was supported by JSPS KAKENHI Grant Number JP24K14897.

% \subsubsection{\discintname}
% The authors have no competing interests to declare that are relevant to the content of this
% article.
\end{credits}

%% -----------------------------------------------------------------------
%% Bibliography
%% -----------------------------------------------------------------------

\bibliographystyle{splncs04}
\bibliography{references}

\end{document}
