%% Section 1: Introduction
%% Japanese draft (draft2/01_introduction.md) embedded as comments.
%% English translation written below each paragraph.
%% -----------------------------------------------------------------------

\section{Introduction}
\label{sec:introduction}

% 整数数列のオンライン百科事典 OEIS（On-Line Encyclopedia of Integer Sequences）[cite:sloane1996] は、
% 2026年1月時点で組合せ論・数論・代数学など数学の幅広い分野にまたがる 391,000 件以上の数列エントリを
% 収録した事実上の標準リファレンスである。
% 各エントリは整数の有限列とその数学的定義を対応付けており、OEIS は機械可読な数学的知識の体系として
% 他に類を見ない存在である。
The On-Line Encyclopedia of Integer Sequences (OEIS)~\cite{sloane1996}, as of January 2026,
catalogues more than 391,000 entries spanning combinatorics, number theory, algebra, and many
other branches of mathematics, making it the \emph{de facto} standard reference for integer
sequences. Each entry associates a finite integer sequence with its mathematical definition,
rendering the OEIS a uniquely machine-readable corpus of mathematical knowledge.

% 本研究が定式化するタスクは マスク付き系列モデリング（masked sequence modelling; 穴埋め）である。
% 数列の複数の位置をランダムにマスクし、モデルは周囲の整数の並びから各マスク位置の値を予測するよう
% 学習される。このタスクは「モデルが整数数列を支配する算術的・組合せ論的法則をどの程度内在化できて
% いるか」を問う直接的な試金石であり、次項予測（next-term prediction）はその特殊ケースとして
% Solver コンポーネントを介した応用評価として位置づける。
% 整数をトークン ID として直接埋め込む先行研究 FACT [cite:zurich-fact] は，穴埋めを含む 6 つの
% タスクを定義・評価し，素の Transformer でどこまで達成できるかの基準を示したが，学習時に未見の
% 大きな整数値への対応と乗法的構造の習得に根本的な限界を抱えている。
The task we formalise is \emph{masked sequence modelling}: a random subset of positions in a
sequence is masked and the model is trained to predict each masked value from its surrounding
context. This task directly probes how well a model has internalised the arithmetic and
combinatorial laws governing integer sequences; next-term prediction is treated as a special
case and evaluated via a dedicated Solver component. The prior benchmark FACT~\cite{zurich-fact}
defines and evaluates six tasks including unmasking, establishing what a plain Transformer can
achieve, but it faces fundamental limitations in handling large integer values outside its fixed
token vocabulary and in learning multiplicative structure.

% このタスクの主要な困難は、整数数列の極端な不均質性にある。
% 値の範囲は一桁の定数から天文学的な階乗・指数関数的数列まで及び、単一の学習コーパス内で数十桁以上の
% 桁数差が存在する。同時に、多くの数列は素数値数列・組合せパリティ・周期的剰余など、値の大小とは
% 独立した剰余制約に従う。
% 各整数を離散語彙トークンに割り当てる標準的なトークン化アプローチは、このような状況に根本的に不向きで
% ある。学習時に未見の大きな整数値を扱えず、算術的構造を不透明なトークン ID に押し込め、大きな数に
% 対してスケールが破綻する。
The principal difficulty of this task is the extreme heterogeneity of integer sequences.
Values range from single-digit constants to astronomically large factorials and exponential
sequences, with differences of tens of orders of magnitude within a single training corpus.
At the same time, many sequences obey \emph{residue constraints}---parity, combinatorial
patterns, or periodic remainders---that are independent of the magnitude of individual values.
The standard tokenisation approach of assigning each integer a discrete vocabulary token is
fundamentally ill-suited to this setting: it cannot represent integers outside its fixed token
vocabulary, embeds arithmetic structure in opaque token IDs, and breaks down in scale for large numbers.

% 本研究では IntSeqBERT を提案する。
% これは OEIS コーパス全体で穴埋め問題により大規模学習（事前学習）された Transformer エンコーダで
% あり、双ストリーム入力表現によって上記の課題を克服する。
% 整数をトークン化するのではなく、IntSeqBERT は各要素を 2 つの相補的な軸で符号化する：
% - Magnitude ストリーム: 対数スケールで絶対値を表現する連続埋め込み。成長挙動とスケールを捉える。
% - Modulo ストリーム: 法 2 から 101 まで 100 個の剰余に対する Sin/Cos 埋め込み。周期性と数論的
%   構造を捉える。
% 2 つのストリームは FiLM（Feature-wise Linear Modulation）[cite:perez2018film] で融合される。
% 学習では、Magnitude 回帰・符号分類・100 個の法に対する Modulo 予測の 3 つの予測器を多目的に
% 同時学習する（マルチタスク事前学習）。マスク位置の予測値（大きさ・符号・剰余分布）から具体的な整数値
% を復元するために、中国剰余定理（CRT）に基づく IntegerSolver を用いる（第 3.9 節）。
We propose \textbf{IntSeqBERT}, a Transformer encoder pre-trained on the OEIS corpus via masked
sequence modelling, which overcomes the above challenges through a \emph{dual-stream} input
representation. Rather than tokenising integers, IntSeqBERT encodes each element along two
complementary axes:
\begin{itemize}
  \item \textbf{Magnitude stream}: a continuous log\-scale embedding of the
        absolute value, capturing growth behaviour and scale.
  \item \textbf{Modulo stream}: sin/cos embeddings for 100 residues modulo $2$ through $101$,
        capturing periodicity and number-theoretic structure.
\end{itemize}
The two streams are fused via FiLM (Feature-wise Linear Modulation)~\cite{perez2018film}.
During training, three predictors are jointly optimised in a multi-task objective: magnitude
regression, sign classification, and modulo prediction for 100 moduli. Concrete integer values
are recovered from the masked-position predictions (magnitude, sign, residue distributions)
using a probabilistic Chinese Remainder Theorem (CRT)-based \textbf{IntegerSolver} (Section~\ref{sec:solver}).

% 本研究では OEIS から抽出した274,705件のデータセットに対して、3 モデルサイズ（Small: ~9M・
% Middle: ~44M・Large: ~110M パラメータ）において IntSeqBERT を 2 つのベースラインと比較した。
% 全実験は GeForce RTX 3070 Ti（VRAM 8 GB）1 枚で実施しており、コンシューマー GPU のメモリ制約下
% での結果である点は解釈の際に留意されたい。
% 主要な発見は以下の通りである：
We evaluate IntSeqBERT against two baselines on 274,705 OEIS sequences across three model
sizes (Small: 6.4M, Middle: 29.0M, Large: 91.5M parameters), all on a
single GeForce RTX 3070 Ti (8\,GB VRAM).

% 貢献のまとめ。
\noindent\textbf{Contributions.}
\begin{enumerate}
  % IntSeqBERT: Magnitude と剰余算術の特徴埋め込みを FiLM で融合した整数数列向け双ストリーム
  % Transformer アーキテクチャ。
  \item \textit{IntSeqBERT architecture}: a dual-stream Transformer fusing magnitude and
        modular-arithmetic embeddings via FiLM, jointly trained on OEIS with magnitude
        regression, sign classification, and 100-dimensional modulo prediction. At the
        Large scale, IntSeqBERT achieves \textbf{95.85\%} magnitude accuracy and
        \textbf{50.38\%} MMA, surpassing Vanilla by $+8.9$\,pt and $+4.5$\,pt, respectively.
        The dual-stream representation yields a \textbf{7.4-fold} improvement in
        Solver-based next-term prediction (Top-1: 19.09\% vs.\ 2.59\%).

  % Modulo スペクトル解析: NIG と φ(m)/m の強い負の相関の発見
  \item \textit{Number-theoretic finding}: modulo spectrum analysis reveals that NIG is
        strongly negatively correlated with $\varphi(m)/m$ (where $\varphi$ denotes Euler's totient function; $r = {-}0.851$, $p < 10^{-28}$),
        providing quantitative evidence that composite moduli aggregate arithmetic structure
        more efficiently via CRT.

  % スケーリング分析
  \item \textit{Scaling behaviour}: modulo accuracy and Solver accuracy improve more steeply
        with model size than magnitude accuracy, suggesting arithmetic reasoning benefits
        disproportionately from increased capacity.
\end{enumerate}

% 論文の構成。
\noindent\textbf{Paper organisation.}
Sections~\ref{sec:background}--\ref{sec:setup} review related work, present the architecture,
and describe the experimental setup. Section~\ref{sec:experiments} reports results, Section~\ref{sec:analysis} provides analyses, and Section~\ref{sec:conclusion} concludes.
